---
title: "A1"
author: "Haoxiang Wang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 4
Load the data
```{r}
library(readxl)
library(tidyverse)
# install.packages("jrvFinance")
library(jrvFinance)
bond_data <- read_excel("bond_data.xlsx")
```

The format here is compounding semi-annually (1 period = 6 months)
```{r}
yield_calcualte <- function(coupon, coupon_n, price){
  cf<-rep(coupon,coupon_n)
  cf[length(cf)]<-cf[length(cf)]+100
  cf<-c(-price,cf)
  #print(cf)
  cf_t<-seq(1, coupon_n)
  #the first coupon is one month and a half from t=0
  cf_t<-cf_t-((6-1.5)/6)
  cf_t<-c(0,cf_t)
  #print(cf_t)
  return(irr(cf=cf, cf.t=cf_t ))
  
}
```

Store the results
```{r}
yield_result <- data.frame(matrix(0, nrow = nrow(bond_data), ncol = 11))
colnames(yield_result) <- c("Time", "1-08", "1-09", "1-10", "1-11", "1-12", "1-15", "1-16", "1-17", "1-18", "1-19")
yield_result[,1]<-(bond_data$coupon_number-((6-1.5)/6))/2
```

Calculate the yield curve
```{r}
for (i in c(1:nrow(yield_result))) {
  for (j in c(1:10)) {
    #since the calculation takes half a year as one period, we need to times the result by 2 to transform it into one year's format
    yield_result[i,j+1]=2*yield_calcualte(coupon = as.numeric(bond_data[i,1])/2,coupon_n = as.numeric(bond_data[i,15]),price = as.numeric(bond_data[i,j+4]))
  }
}
```

Plot the yield curve
```{r}
yield_result_long <- pivot_longer(yield_result, -Time, names_to = "Day", values_to = "Yield")
ggplot(yield_result_long, aes(x = Time, y = Yield, color = Day)) +
  geom_line() +
  labs(x = "Time to Maturity", y = "Yield", title = "Yield Curve") +
  theme_minimal() +
  theme(legend.position = "top")
```

Define the spot curve function
```{r}
bootstrap_yield_curve <- function(price, coupon_rate, maturity) {
  n_bonds <- length(price)
  spot_rates <- numeric(n_bonds)
  
  for (i in seq_along(price)) {
    if (i == 1) {
      spot_rate <- ((100 + coupon_rate[i]) / price[i])^(1 / maturity[i]) - 1
      spot_rates[i] <- spot_rate
      
    } else {
      # Calculate the cash flow before the final payment
      cash_flows <- c(rep(coupon_rate[i], i-1))
      discounted_cash_flows <- sapply(1:length(cash_flows), function(j) cash_flows[j] / (1 + spot_rates[j])^maturity[j])
      residual <- price[i] - sum(discounted_cash_flows)
      spot_rates[i] <- ((100 + coupon_rate[i]) / residual)^(1 / maturity[i]) - 1
    }
  }
  return(spot_rates)
}
```

Calculate the spot curve
```{r}
maturity_spot<-seq(1, 11)-((6-1.5)/6)
coupon_rate_spot<-as.numeric(bond_data$coupon)/2
spot_result<-yield_result

for (i in c(1:10)){
  price_spot<-unlist(bond_data[,4+i])
  spot_cal=bootstrap_yield_curve(price_spot, coupon_rate_spot, maturity_spot)
  spot_result[,i+1]=spot_cal*2
}
```

Plot the spot curve
```{r}
spot_result_long <- pivot_longer(spot_result, -Time, names_to = "Day", values_to = "Spot")
ggplot(spot_result_long, aes(x = Time, y = Spot, color = Day)) +
  geom_line() +
  labs(x = "Time to Maturity", y = "Spot Rate", title = "Spot Curve") +
  theme_minimal() +
  theme(legend.position = "top")
```

Calculate the forward curve
```{r}
forward_result<-yield_result[1:7,]
#forward_result[,1]<-c("1yr-1yr","1yr-1.5yr","1yr-2yr","1yr-2.5yr","1yr-3yr","1yr-3.5yr","1yr-4yr")
forward_result[,1]<-c(1,1.5,2,2.5,3,3.5,4)
for (i in c(1:10)) {
  for (j in c(1:7)) {
    f_1<-(1+(spot_result[j+4,i+1]/2))^(spot_result[j+4,1]*2)
    #print(spot_result[j+4,i+1])
    #print(spot_result[j+4,1]*2)
    f_2<-(1+spot_result[3,i+1]/2)^(spot_result[3,1]*2)
    #print(spot_result[3,i+1])
    #print(spot_result[3,1]*2)
    f_3<-1/(spot_result[j+4,1]*2-spot_result[3,1]*2)
    forward_result[j,i+1]<-((f_1/f_2)^(f_3)-1)*2
  }
}

```

Plot the forward curve
```{r}
forward_result_long <- pivot_longer(forward_result, -Time, names_to = "Day", values_to = "Spot")
ggplot(forward_result_long, aes(x = Time, y = Spot, color = Day)) +
  geom_line() +
  labs(x = "Forward Time (from Year 1)", y = "Forward Rate", title = "Forward Curve") +
  theme_minimal() +
  theme(legend.position = "top")
```

## 5
Consider the yield first
```{r}
yield_Q5<-yield_result[, -1]
yield_Q5<-yield_Q5[c(3, 5, 7, 9, 11), ]
x_yield<-matrix(1:45, nrow = 5)
for (i in c(1:9)) {
  x_yield[,i]<-log(yield_Q5[,i+1]/yield_Q5[,i])
}

cov_yield<-cov(t(x_yield))
cov_yield
```

Now, consider the forward
```{r}
forward_Q5<-forward_result[,-1]
forward_Q5<-forward_Q5[c(1, 3, 5, 7), ]
x_forward<-matrix(1:36, nrow = 4)
for (i in c(1:9)) {
  x_forward[,i]<-log(forward_Q5[,i+1]/forward_Q5[,i])
}

cov_forward<-cov(t(x_forward))
cov_forward
```

## 6
Yield's covariance matrix
```{r}
eigen_result_y <- eigen(cov_yield)
eigenvalues_y <- eigen_result_y$values
eigenvectors_y <- eigen_result_y$vectors
print("Eigenvalue")
print(eigenvalues_y)
print("Eigenvector")
print(eigenvectors_y)
```

Forward's covariance matrix
```{r}
eigen_result_f <- eigen(cov_forward)
eigenvalues_f <- eigen_result_f$values
eigenvectors_f <- eigen_result_f$vectors
print("Eigenvalue")
print(eigenvalues_f)
print("Eigenvector")
print(eigenvectors_f)
```




